---
title: "Module 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains all of the R coding we will use in Module 2.


## 1.1 Working Directories

<!-- setwd("/Users/pkvam/Desktop/Rprograms/209") -->

```{r}
getwd()
```
Before you start generating your own files, set your working directory to where you want you output to be stored (and to where you want R to look for your input files) using the R command `setwd()`


## 1.2   Libraries

```{r}
library(MASS)
```

## 1.3 Imported Data
```{r}
summary(Rabbit)
```


```{r}
x1 <- c(6,1,30,9,12,2,21,5)
sample(x1,replace=TRUE)
```

```{r}
head(Rabbit,n=10)
```


## 1.4 R Commands

```{r}
4*5 + 2^3 - exp(1)
```
```{r}
sqrt(2)
```

```{r}
rep(2,10)
```

#### Lists and Vectors

```{r}
rep(c(1,2,3),4)
```

```{r}
rep(c("a",31),4)
```

```{r}
seq(from=10, to=40, by=5)
```

```{r}
x1 <- c(6,1,30,9,12,2,21,5)
x1
```

```{r}
length(x1)
```

```{r}
x1[4]
```

```{r}
x1[-1]
```

#### Ranking and sorting

```{r}
sort(x1)
```

```{r}
rank(x1)
```

```{r}
c(min(x1),mean(x1),median(x1),sd(x1),max(x1))
```

#### Logical Operators

```{r}
x1 == 30
```

```{r}
(x1 == 30) + 0
```

```{r}
sum(x1<10)
```

```{r}
rev(x1)
```

```{r}
which(x1>10)
```

```{r}
x1[which(x1>10)]
```

```{r}
x1[x1>10]
```

```{r}
x1[x1!=9]
```

#### Vector Operations

```{r}
x1*10 
```

```{r}
x1^2 
```


```{r}
x1*x1 
```

#### Advanced Operations

```{r}
sample(x1)
```

```{r}
sample(x1,replace=TRUE)
```

```{r}
range(x1)
```

```{r}
x1[[9]]<-50
x1
```

```{r}
matrix(1:12,nrow=3) 
```



## 2 Dr. Arbuthnot’s Baptism Records

```{R}
source("http://www.openintro.org/stat/data/arbuthnot.R")
head(arbuthnot,10)
names(arbuthnot)
arbuthnot[1,]
```

```{R}
hist(arbuthnot$boys)
```

```{R}
plot(x = arbuthnot$year , y = arbuthnot$girls)
```

```{R}
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l" )
```

```{R}
arbuthnot$boys + arbuthnot$girls
```

```{R}
plot(arbuthnot$year, arbuthnot$boys + arbuthnot$girls , type = "l" )
```

```{R}
arbuthnot$boys /( arbuthnot$boys + arbuthnot$girls )
```

```{R}
arbuthnot$boys[arbuthnot$year>1699]
```




#### end of module 2
